\selectlanguage{british}
\chapter*{Abstract}
\markboth{Abstract}{Abstract}
Structural topology optimization is now a well-developed tool that many engineers can easily use to determine optimal structures for minimal weight. Thermal topology optimization is also becoming a well-studied subject with many methodologies now being developed. By integrating these tools into a single module, in theory, a structure can be found that optimizes both problems. Usually, these induce some trade-off between the single problems that need to be considered. The idea of Pareto optimality arises when a structure can be found to minimize the error of the structure with respect to some unobtainable design criteria. 

Of course, material properties need to be known to optimize the domain. Analytical and semi-analytical models have been developed for many lattice unit cells. These models will be utilized in this thesis.

In this thesis, a methodology for multi-functional topology optimization has been developed as well as a basic implementation written in the Julia programming language. The finite element method has been utilized for resolving structural, thermal and phase change problems. An optimal structure can be determined given a so-called Utopia point which generally can not be found exactly. Tools for solving single-functional topology optimization problems have also been implemented that consider the homogenized properties of selected lattice unit cells.

\textbf{Keywords:} lattice structures, phase change materials, multi-functional, topology optimization, Pareto optimality