\chapter{Source Code}

\section*{Main Files}
The following files are the main files.

\subsection*{main.jl}
This is the main Julia file. This file acts as a .ini file that loads all the required data. It imports a mesh and preallocates some matrices that are required for the topology optimization functions.
\lstinputlisting{../src/main.jl}

\subsection*{phaseChange.jl}
This file has the phase change methods for solving the phase change problem.
\lstinputlisting{../src/phaseChange.jl}

\subsection*{optimizeFunctions.jl}
This file has all the main optimization functions and solve functions. This file can generate the Pareto fronts, and structures for the different topology optimization cases.
\lstinputlisting{../src/optimiseFunctions.jl}

\subsection*{topologyOptimisation.jl}
This file has the main topology optimization algorithms such as the filtering and sensitivity analysis methods.
\lstinputlisting{../src/topologyOptimisation.jl}

\subsection*{latticeProperties.jl}
This file has all the lattice properties and related methods. This is also where the lattice material properties are defined.
\lstinputlisting{../src/latticeProperties.jl}


\section*{Helper Files}
The following files are helper files for things such as shape functions and file io.

\subsection*{shapeFunctions.jl}
This file defines the shape functions and related methods.
\lstinputlisting{../src/shapeFunctions.jl}

\subsection*{assembly.jl}
This file handles all the finite element assembly, as well as global system assembly.
\lstinputlisting{../src/assembly.jl}

\subsection*{fileio.jl}
This file handles all the file input/output, such as importing the mesh and exporting results to .vtu or .pvd files.
\lstinputlisting{../src/fileio.jl}